\section{Conclusion}

In this paper, we have focused on the retrievability and processability
 of the WOD for machine agents.
We have therefore not focused on the interactions between datasets,
 i.e. the fifth star of LOD publishing.
We would like to extend \obs~to also take relations between individual
 resources into account.
Part of this effort would consist of redrawing the LOD cloud,
 but for more datasets than this was ever done in the past.

Another extension of our current research is to run the \obs
 script on more CKAN datastores.
%Many resources in CKAN have an e-mail address of the resource's maintainer.
%Our implementation is able to automatically send an email to these address
% once a problem with the associated resource is encountered.
%We would like to use this automated feature in the future to stimulate
% dataset maintainers to make their data fully machine-findable
% and -processable.
Since the implemented script can run in a fully automated fashion,
 we would also like to rerun it on a regular basis and have \obs~
 keep track of the (almost) real-time availability
 of data on the Web.
%We are currently waiting for our faculty to obtain an application server
% on which the observatory will be hosted.

In summary, we have provided a detailed operationalization of
 the 5-star publication model for Linked Data,
 which we have built into a Web Observatory, called \obs.
This observatory differs from existing ones, as it focuses on monitoring
 the machine friendliness of data.
We show that in its current state, much of the Linked Data available on
 the Web is far from reaching the 5-star level.
This is not just a technical issue but a social issue
 where the dynamics of the Web's social technical system have not reached
 a point where machine friendly data is widely available.
By providing an observatory on the state of the machine processability of
 Web data, we hope to guide interventions at both the technical
 and social level.
Additionally, this observatory will help in tracking the outcome of
 those interventions.

